A Discontinuous Petrov-Galerkin method for convection-diffusion problems

### 2013-07-30 docs/Jesse/Dissertation/dissPres.tex

\title[A DPG method for Compressible Flow]{A Discontinuous Petrov-Galerkin method for compressible flow problems}
\date{August 8, 2012}
\author[J. Chan, L. Demkowicz, R. Moser]{Jesse Chan \\ Supervisors: Leszek Demkowicz, Robert Moser \\ Committee: Todd Arbogast, Omar Ghattas, Venkat Raman}

@@

\section{Introduction/Literature review}
\frame{
\frametitle{Phenomena in compressible flow}
\begin{figure}
\centering
\subfigure[Shock wave]{\includegraphics[scale=.0704]{figs/compFlowPics/bullet_shock.png}}
\subfigure[Boundary layer]{\includegraphics[scale=.16]{figs/compFlowPics/boundary_layer.png}}
%\subfigure{\includegraphics[scale=.15]{figs/compFlowPics/capsule.png}}
\end{figure}
Compressible flow plays an important role in the aerospace and energy industries - transonic and supersonic aircraft, combustion engines, etc.

}
\frame{
\frametitle{Compressible Navier-Stokes equations}
\begin{columns}[c]
\begin{column}{.475\textwidth}
Numerical difficulties:
\begin{itemize}
\item{} Resolving solution features (sharp, localized viscous-scale phenomena)
\begin{itemize}
\item{} Shocks
\item{} Boundary layers - resolution needed for drag/load
\item{} Turbulence (non-localized)
\end{itemize}
\item{} Nonlinear convergence and uniqueness of solutions
\item{} Stability of numerical schemes
\begin{itemize}
\item{} Coarse/adaptive grids
\item{} Higher order
\end{itemize}
\end{itemize}
\end{column}
\begin{column}{.475\textwidth}
\begin{figure}
\centering
\includegraphics[scale = .7]{figs/compFlowPics/airfoil_mesh.png}
\end{figure}
\end{column}
\end{columns}
\vspace{.2cm}
\begin{center}
Idea: begin first with the model problem of convection-diffusion.
\end{center}
}
\frame{
\frametitle{Robustness: convection-diffusion as a model problem}
\vspace{-.5cm}
\[
\div \left(\beta u\right) - \epsilon \Delta u = f, \quad \text{on }\Omega \in \mathbb{R}^3
\]
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.49\textwidth}
In 1D: $\beta u' - \epsilon u'' = f$. Standard continuous Galerkin variational formulation: solve
\[
b(u,v) = \ell(v), \quad u,v\in H^1_0([0,1])
\]
where
\begin{align*}
b(u,v) &= \int_\Omega -\beta uv' + \epsilon u'v' \\
\ell(v) &= \int_\Omega f v 
\end{align*}
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.35]{figs/GalerkinOscTight.png}
\caption{Solution for $f=1$. Oscillations in the standard Galerkin method for large Peclet numbers ${\rm Pe} \coloneqq \frac{h}{\epsilon}$.}
\end{figure}
\end{column}
\end{columns}
}

@@

\frame{
\begin{columns}
\begin{column}{.49\textwidth}
Can be interpreted as a \emph{Petrov-Galerkin} method,
\[
b\LRp{u,\tilde{v}_{i}} = \ell\LRp{\tilde{v}_{i}}, \quad \forall i = 1,\ldots,N-1,
\]
where the SUPG test function $\tilde{v}_{i}$ is defined elementwise as\footnotemark
\[
\tilde{v}_{i} = \phi_i(x) + \tau L_{\rm adv} \phi_i.  
\]
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale=.16]{figs/SUPGtest.png}}
\caption{SUPG test function $v_i$.}
\end{figure}
\end{column}
\end{columns}
\footnotetext{\bibentry{SUPG}}
}

\section{The DPG Method}
\frame{
\frametitle{DPG: a minimum residual method via optimal testing}
Given a trial space $U$ and Hilbert test space $V$,
\[
b(u,v) = \ell(v), \quad u\in U, \quad v\in V
\]
This is equivalent to the operator equation posed in $V'$
\[
Bu = \ell
\]
if we identify $B:U\rightarrow V'$ and $\ell\in V'$ such that 
\begin{align*}
\langle Bu,v\rangle_V &\coloneqq b(u,v), \quad u\in U, v\in V,\\
\langle \ell,v\rangle_V &\coloneqq \ell(v), \quad v\in V.
\end{align*}
We seek to minimize the dual functional over $U_h \subset U$
\[
J(u_h) = \frac{1}{2}\|Bu_h-\ell\|_{V'}^2 \coloneqq\frac{1}{2} \sup_{v\in V\setminus\{0\}} \frac{| b(u_h,v)-\ell(v)|^2}{\nor{v}_V^2}.
\]
}

\begin{frame}
Let $R_V: V \to V'$ be the isometric Riesz map st.\
\[
\langle R_V v,\delta v\rangle_V \coloneqq(v, \delta v)_V, \quad \forall \delta v \in V.
\]
Then, our functional $J(u_h)$ is equal to
\begin{equation*}
\min_{u_h\in U_h} J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2 =  \frac{1}{2}\left\|R_V^{-1}(Bu_h-\ell)\right\|_V^2.
\end{equation*}
First order optimality: G\^ateaux derivative is zero in all directions $\delta u \in U_h$
\begin{align*}
\left(R_V^{-1}(Bu_h-\ell),R_V^{-1}B\delta u\right)_V &= 0, \quad \forall \delta u \in U_h.\\
\rightarrow \LRa{(Bu_h-\ell),R_V^{-1}B\delta u} &= 0, \\
\rightarrow b(u_h,R_V^{-1}B\delta u) - \ell(R_V^{-1}B\delta u) &= 0
\end{align*}
\end{frame}

\frame{
\frametitle{Summary: select test functions to minimize residuals}
For $\delta u \in U_h$, define the \textcolor{red}{optimal test function} $v_{\delta u}$. 
\begin{equation*}
v_{\delta u} \coloneqq R_V^{-1}B\delta u.
\end{equation*} 
Then, the residual $J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2$ is minimized by the solution of
\[
b\LRp{u_h,v_{\delta u}} = \ell\LRp{v_{\delta u}}, \quad \forall \delta u \in U_h.
\]
}

\frame{
\frametitle{Practical details of DPG}
Computation of $v_{\delta u} \coloneqq R_V^{-1}B\delta u$ is \textbf{global} and \textbf{infinite-dimensional}.
\begin{itemize}
\item By choosing a \textcolor{red}{broken} test space $V$ and \textcolor{red}{localizable} norm $\nor{v}_V^2 = \sum_K \nor{v}_{V(K)}^2$, test functions can be determined locally.
\item In practice, we use an \textcolor{red}{enriched space} $V_h \subset V$, where $\dim(V_h) > \dim(U_h)$ elementwise, and \textcolor{red}{optimal test functions} are approximated by computing $v_{\delta u} \coloneqq R_{V_h}^{-1}B\delta u \in V_h$ through\footnote{\bibentry{DPG2}}
\[
\LRp{v_{\delta u},\delta v}_V = b(\delta u,\delta v), \quad \delta u\in U_h, \quad \forall \delta v\in V_h
\]
Typically, if $U_h = \mathcal{P}^p(\mathbb{R}^n)$, $V_h = \mathcal{P}^{p+\triangle p}(\mathbb{R}^n)$, where $\triangle p \geq n$.\footnote{\bibentry{practicalDPG}} 
\end{itemize}
%\footnotetext{\bibentry{DPG2}}
}

@@

\frame{
\frametitle{Properties of DPG}
Additional properties of DPG include\footnotemark
\begin{itemize}
\item DPG provides the best approximation in the \textcolor{red}{energy norm}
\[
\|u\|_E = \|Bu\|_{V'} = \sup_{\nor{v}_V=1} \left|b(u,v)\right|.
\]
\item The energy error is computable through the residual
\[
\nor{u-u_h}_E = \nor{B(u-u_h)}_{V'} = \nor{R_V^{-1}(l-Bu_h)}_V = \nor{e}_V
\]
where the \textcolor{red}{error representation function} $e$ is defined through $(e,\delta v)_V = \ell(v)-b(u_h,\delta v)$ for all $\delta v\in V$. 
\end{itemize}
\footnotetext{\bibentry{DPG3pub}}
}

%\frame{
%\frametitle{Duality of norms}
%Duality of norms: for any trial norm $\nor{u}_U$,
%\[
%\nor{u}_{U} = \sup_{v \in V}\frac{b\LRp{u,v}}{\nor{v}_{V,U}}, \quad \nor{v}_{V,U} = \sup_{u \in U}\frac%{b\LRp{u,v}}{\nor{u}_{U}}.
%\]
%Likewise, for any test norm $\nor{v}_V$, 
%\[
%\nor{v}_{V} = \sup_{u \in U}\frac{b\LRp{u,v}}{\nor{U}_{U,V}}, \quad \nor{u}_{U,V} = \sup_{v \in V}\frac%{b\LRp{u,v}}{\nor{v}_{V}}.
%\]
%
%}

\frame{
\frametitle{Ultra-weak formulation}
\vspace{-.75cm}
\begin{columns}
\begin{column}{.64\textwidth}
Given a first order system $Au = f$, we identify the \textbf{\emph{partition}} $\Oh$ and \textbf{\textcolor{red}{mesh skeleton}} \textcolor{red}{$\Gh$}.
\end{column}
\begin{column}{.34\textwidth}
%\[
%\Oh = \cup_{j=1}^\Nel K_j, \quad \Gh = \cup_{j=1}^\Nel \partial K_j
%\]
\begin{figure}[!h]
\centering
\includegraphics[scale = .155]{mesh_skel.pdf}
\end{figure}
\end{column}
\end{columns}
\vspace{-.5cm}
The ultra-weak formulation for $Au = f$ on $\Oh$ is
\[
b\left(\left(u, \widehat{u}\right),v\right) \coloneqq \sum_K\langle \widehat{u},v \rangle_{\partial K} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh}.
\]
Under proper assumptions, $\sum_K\langle \widehat{u},v \rangle_{\partial K} = \langle \widehat{u}, \jump{v} \rangle_{\Gh}$, with energy setting
\[
u\in L^2\LRp{\Oh} \equiv L^2(\Omega), \quad v\in V=D(A^*_h), \quad
\widehat{u}\in \gamma(D(A)),
\]
where $D(A_h^*)$ is the broken graph space of the formal adjoint $A_h^*$, and $\gamma(D(A))$ the trace space of the graph space of operator $A$.\footnote{\bibentry{analysisDPG}}
}

\frame{
\frametitle{The canonical ``graph'' test norm}
Recall $\nor{u}_E \coloneqq \sup_{v\in V\setminus\{0\}}\frac{b(u,v)}{\nor{v}_V}$. Under the ultra-weak formulation, the trial norm 
\[
\nor{\LRp{u,\widehat{u}}}_U \coloneqq \|u\|^2_{\L} + \|\widehat{u}\|^2
\] 
generates a test norm \emph{equivalent} to 
\[
\nor{v}_{V_{\rm opt}} \coloneqq \|A_h^*v\|_{\L}^2 +\left(\sup_{\widehat{u}} \frac{\LRa{ \widehat{u},\jump{v} }_{\Gh}}{\|\widehat{u}\|}\right)^2.
\]
This norm is not localizable, so we instead substitute the jump terms for an $L^2$ term, giving us the \emph{graph norm}\footnote{\bibentry{Bui-ThanhDemkowiczGhattas11b}}
\[
\nor{v}_V \coloneqq \|A_h^*v\|_{\L}^2 + \nor{v}_{L^2}.
\]
}

\frame{
\frametitle{Goal: robust higher order adaptive methods}
\begin{columns}
\begin{column}{.45\textwidth}
Our aim is to design an adaptive method for compressible laminar flows that is 
\begin{itemize}
\item robust in Reynolds number
\item stable for arbitrary elements
\end{itemize}

This proposal will outline
\begin{enumerate}
\item a robust DPG method for convection-diffusion,
\item extension to nonlinear problems and systems,
\item proposed work.
\end{enumerate}

\end{column}
\begin{column}{.54\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale=.23]{figs/Re1000p2/u2_full.png}}
\subfigure{\includegraphics[scale=.37]{figs/Re1000p2/mesh6.png}}
\caption{$u_2$ and $p=2$ mesh for $\Reyn = 1000$.}
\end{figure}
\end{column}
\end{columns}
}

\section{A Robust DPG Method for Convection-Diffusion}

\frame{
\frametitle{Ultra-weak formulation for convection-diffusion}
In first order form, the convection-diffusion equation is
\[
A \LRp{u,\sigma} \coloneqq 
\LRs {
\begin{array}{c}
\div (\beta u - \sigma) \\ \frac{1}{\epsilon}\sigma - \grad u
\end{array}} = \LRs{
\begin{array}{c}
f \\ 0
\end{array}
}.
\]
The variational formulation is
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\grad_h\cdot \tau - \beta \cdot \grad_h
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad_h v\right)_{\Oh}&\\
  - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}&,
\end{align*}
where $\widehat{f}_n \coloneqq \beta_n u - \sigma_n$ and $\LRa{\widehat{f}_n,\jump{v} }_{\Gh}$ is defined 
\[
\LRa{\widehat{f}_n,\jump{v} }_{\Gh}\coloneqq \sum_K\int_{\partial K}{\rm sgn}(\vec{n}) \, \widehat{f}_n v.
\]
}

\frame{
\frametitle{Graph norm under convection-diffusion}
For convection-diffusion, the graph test norm is defined elementwise
\[
\|\left(v,\tau\right)\|_{V(K)}^2 = \| \div \tau - \beta \cdot \grad v
\|_{L^2(K)}^2 + \| \epsilon^{-1} \tau + \grad v \|_{L^2(K)}^2 +
\|v\|_{L^2(K)}^2.
\]
Problem with this test norm: approximability of test functions.
\vspace{-.3cm}
\begin{figure}[!h]
\centering
\includegraphics[scale=.175]{figs/opt.png}
\caption{$v$ and $\tau$ components of the 1D optimal test functions for flux $\widehat{f}_n$ on the \emph{right-hand} side of a unit element for $\epsilon = 0.01$. }
\label{fig:optTestBoundary}
\end{figure}
}

\frame{
\frametitle{Determining an alternative test norm}
Recall the convection-diffusion bilinear form
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\div \tau - \beta \cdot \grad
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad v\right)_{\Oh}&\\
 - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}&,
\end{align*}
We recover $\nor{u}^2_{L^2(\Omega)}$ by choosing conforming $\LRp{v,\tau}$ satisfying the \emph{adjoint equations}
\begin{align*}
\div\tau - \beta \cdot \grad v   &= u\\
\frac{1}{\epsilon}\tau + \grad v &= 0
\end{align*}
with boundary conditions s.t.\ $\LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gamma}$ and $\LRa{ \widehat{f}_n,\jump{v} }_{\Gamma}$ vanish.
}

\frame{
\frametitle{A robust bound}
``Necessary'' conditions for robustness --- let $\boldsymbol U = \LRp{u,\sigma,\widehat{u},\widehat{f}_n}$. Then, by choosing specific $\LRp{v,\tau}$ satisfying the adjoint equations,
\[
\nor{u}^2_{L^2(\Omega)} = b\LRp{{\boldsymbol U},\LRp{v,\tau}} = \frac{b\LRp{{\boldsymbol U},\LRp{v,\tau}}}{\nor{\LRp{v,\tau}}_V} \nor{\LRp{v,\tau}}_V \leq \nor{\boldsymbol U}_E \nor{\LRp{v,\tau}}_V
\]
Let $\lesssim$ denote a robust bound - \textcolor{red}{if $ \nor{\LRp{v,\tau}}_V \lesssim \|u\|_{L^2(\Omega)}$}, then we have that
\[
\nor{u}_{L^2(\Omega)} \lesssim \nor{\boldsymbol U}_E
\]
\textbf{Main idea: the \textcolor{red}{test norm} should measure \textcolor{red}{adjoint solutions} robustly.}
}

>>>>

\title[DPG for convection-diffusion]{A Discontinuous Petrov-Galerkin method for convection-diffusion problems}
\date{July 17, 2013}
\author[J. Chan, L. Demkowicz, R. Moser]{Jesse Chan \\ Supervisors: Leszek Demkowicz, Robert Moser \\ Committee: Todd Arbogast, Omar Ghattas, Venkat Raman}

@@

\begin{frame}
\frametitle{Acknowledgements}
\begin{columns}[c]
\begin{column}{.03\textwidth}
\end{column}
\begin{column}{.33\textwidth}

Omar Al-Hinai

Ivo Babu\v{s}ka

Paul Baumann

Tan Bui

Henry Chang

Truman Ellis

John Evans

\end{column}
\begin{column}{.33\textwidth}

Jay Gopalakrishnan

Courtney Grymonprez

John Hawkins

Norbert Heuer

Jessica Meixner

Steve Mattis

Todd Oliver

\end{column}
\begin{column}{.33\textwidth}

Weifeng Qiu

Nate Roberts

Karl Schulz

Chris Simmons

Roy Stogner

Paul Tsuji

Rhys Ulerich

\end{column}

\end{columns}
\vspace{1cm}
\begin{center}
\ldots and many more.
\end{center}
\end{frame}

\section{Introduction/Literature review}
\frame{
\frametitle{Phenomena in compressible flow}
\begin{figure}
\centering
\subfigure[Shock wave]{\includegraphics[scale=.0704]{figs/compFlowPics/bullet_shock.png}}
\subfigure[Boundary layer]{\includegraphics[scale=.16]{figs/compFlowPics/boundary_layer.png}}
%\subfigure{\includegraphics[scale=.15]{figs/compFlowPics/capsule.png}}
\end{figure}
Compressible flow plays an important role in the aerospace and energy industries - transonic and supersonic aircraft, combustion engines, etc. 
}
\frame{
\frametitle{Compressible Navier-Stokes equations}
\begin{columns}[c]
\begin{column}{.475\textwidth}
Numerical difficulties:
\begin{itemize}
\item{} Resolving solution features (sharp, localized viscous-scale phenomena)
\begin{itemize}
\item{} Shocks
\item{} Boundary layers
\item{} Turbulence (non-localized)
\end{itemize}
\item{} Nonlinear convergence and uniqueness of solutions
\item{} Stability of numerical schemes
\begin{itemize}
\item{} Coarse/adaptive grids
\item{} Higher order
\end{itemize}
\end{itemize}
\end{column}
\begin{column}{.475\textwidth}
\begin{figure}
\centering
\includegraphics[scale = .7]{figs/compFlowPics/airfoil_mesh.png}
\end{figure}
\end{column}
\end{columns}
\vspace{.2cm}
\begin{center}
Idea: begin first with the model problem of convection-diffusion.
\end{center}
}

\begin{frame}{Outline}
  \tableofcontents[currentsection]
\end{frame}

\frame{
\frametitle{Robustness: convection-diffusion as a model problem}
\vspace{-.5cm}
\[
\div \left(\beta u\right) - \epsilon \Delta u = f, \quad \text{on }\Omega \in \mathbb{R}^3
\]
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.49\textwidth}
In 1D: $\beta u' - \epsilon u'' = f$. Standard continuous Galerkin variational formulation: solve
\[
b(u,v) = \ell(v), \quad u,v\in H^1_0([0,1])
\]
where
\begin{align*}
b(u,v) &= \int_\Omega -\beta uv' + \epsilon u'v' \\
\ell(v) &= \int_\Omega f v 
\end{align*}
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.35]{figs/GalerkinOscTight.png}
\caption{Solution for $f=1$. Oscillations in the Galerkin method for large Peclet numbers ${\rm Pe} \coloneqq \frac{h}{\epsilon}$.}
\end{figure}
\end{column}
\end{columns}
}

@@

\frame{
\begin{columns}
\begin{column}{.49\textwidth}
Can be interpreted as a \emph{Petrov-Galerkin} method,
\[
b\LRp{u,\tilde{v}_{i}} = \ell\LRp{\tilde{v}_{i}}, \quad \forall i = 1,\ldots,N-1,
\]
where the SUPG test function $\tilde{v}_{i}$ is defined elementwise as\footnotemark
\[
\tilde{v}_{i} = \phi_i(x) + \tau L_{\rm adv} \phi_i.  
\]
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale=.16]{figs/SUPGtest.png}}
\caption{SUPG test function $v_i$.}
\end{figure}
\end{column}
\end{columns}
\footnotetext{\tiny\bibentry{SUPG}}
}

\section{The DPG Method}
\begin{frame}{Outline}
  \tableofcontents[currentsection]
\end{frame}

\frame{
\frametitle{DPG: a minimum residual method via optimal testing}
Given a trial space $U$ and Hilbert test space $V$,
\[
b(u,v) = \ell(v), \quad u\in U, \quad v\in V
\]
This is equivalent to the operator equation posed in $V'$
\[
Bu = \ell
\]
if we identify $B:U\rightarrow V'$ and $\ell\in V'$ such that 
\begin{align*}
\langle Bu,v\rangle_V &\coloneqq b(u,v), \quad u\in U, v\in V,\\
\langle \ell,v\rangle_V &\coloneqq \ell(v), \quad v\in V.
\end{align*}
We seek to minimize the dual functional over $U_h \subset U$
\[
J(u_h) = \frac{1}{2}\|Bu_h-\ell\|_{V'}^2 \coloneqq\frac{1}{2} \sup_{v\in V\setminus\{0\}} \frac{| b(u_h,v)-\ell(v)|^2}{\nor{v}_V^2}.
\]
}

\begin{frame}
Let $R_V: V \to V'$ be the Riesz map (an isometry):\
\[
\langle R_V v,\delta v\rangle_V \coloneqq(v, \delta v)_V, \quad \forall \delta v \in V.
\]
Then, our functional $J(u_h)$ is equal to
\begin{equation*}
\min_{u_h\in U_h} J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2 =  \frac{1}{2}\left\|R_V^{-1}(Bu_h-\ell)\right\|_V^2.
\end{equation*}
First order optimality: G\^ateaux derivative is zero in all directions $\delta u \in U_h$
\begin{align*}
\left(R_V^{-1}(Bu_h-\ell),R_V^{-1}B\delta u\right)_V &= 0, \quad \forall \delta u \in U_h.\\
\rightarrow \LRa{(Bu_h-\ell),R_V^{-1}B\delta u} &= 0, \\
\rightarrow b(u_h,R_V^{-1}B\delta u) - \ell(R_V^{-1}B\delta u) &= 0
\end{align*}
Summary: select test functions \textcolor{red}{$v_{\delta u} \coloneqq R_V^{-1}B\delta u$} to \textcolor{red}{minimize residuals}.  
\end{frame}

%\frame{
%\frametitle{Summary: select test functions to minimize residuals}
%For $\delta u \in U_h$, define the \textcolor{red}{optimal test function} $v_{\delta u}$. 
%\begin{equation*}
%v_{\delta u} \coloneqq R_V^{-1}B\delta u.
%\end{equation*} 
%Then, the residual $J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2$ is minimized by the solution of
%\[
%b\LRp{u_h,v_{\delta u}} = \ell\LRp{v_{\delta u}}, \quad \forall \delta u \in U_h.
%\]
%}

\frame{
\frametitle{Practical details of DPG}
Computation of $v_{\delta u} \coloneqq R_V^{-1}B\delta u$ is \textbf{global} and \textbf{infinite-dimensional}.
\begin{itemize}
\item By choosing a \textcolor{red}{broken} test space $V$ and \textcolor{red}{localizable} norm $\nor{v}_V^2 = \sum_K \nor{v}_{V(K)}^2$, test functions can be determined locally.
\item In practice, we use an \textcolor{red}{enriched space} $V_h \subset V$, where $\dim(V_h) > \dim(U_h)$ elementwise, and \textcolor{red}{optimal test functions} are approximated by computing $v_{\delta u} \coloneqq R_{V_h}^{-1}B\delta u \in V_h$ through\footnote{\tiny\bibentry{DPG2}}
\[
\LRp{v_{\delta u},\delta v}_V = b(\delta u,\delta v), \quad \delta u\in U_h, \quad \forall \delta v\in V_h
\]
Typically, if $U_h = \mathcal{P}^p(\mathbb{R}^n)$, $V_h = \mathcal{P}^{p+\triangle p}(\mathbb{R}^n)$, where $\triangle p \geq n$.\footnote{\tiny\bibentry{practicalDPG}} 
\end{itemize}
\let\thefootnote\relax\footnotetext{\tiny\bibentry{DPG2}}
}

@@

\frame{
\frametitle{Properties of DPG}
Additional properties of DPG include\footnotemark
\begin{itemize}
\item DPG provides the best approximation in the \textcolor{red}{energy norm}
\[
\|u\|_E = \|Bu\|_{V'} = \sup_{\nor{v}_V=1} \left|b(u,v)\right|.
\]
\item The energy error is computable through the residual
\[
\nor{u-u_h}_E = \nor{B(u-u_h)}_{V'} = \nor{R_V^{-1}(\ell-Bu_h)}_V = \nor{e}_V
\]
where the \textcolor{red}{error representation function} $e$ is defined through $(e,\delta v)_V = \ell(v)-b(u_h,\delta v)$ for all $\delta v\in V$. 
\end{itemize}
\footnotetext{\tiny\bibentry{DPG3}}
}

\frame{
\frametitle{Ultra-weak formulation}
\vspace{-.75cm}
\begin{columns}
\begin{column}{.64\textwidth}
Given a first order system $Au = f$, we identify the \textbf{\emph{partition}} $\Oh$ and \textbf{\textcolor{red}{mesh skeleton}} \textcolor{red}{$\Gh$}.
\end{column}
\begin{column}{.34\textwidth}
%\[
%\Oh = \cup_{j=1}^\Nel K_j, \quad \Gh = \cup_{j=1}^\Nel \partial K_j
%\]
\begin{figure}[!h]
\centering
\includegraphics[scale = .155]{mesh_skel.pdf}
\end{figure}
\end{column}
\end{columns}
\vspace{-.5cm}
The ultra-weak formulation for $Au = f$ on $\Oh$ is
\[
b\left(\left(u, \widehat{u}\right),v\right) \coloneqq \sum_K\langle \widehat{u},v \rangle_{\partial K} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh}.
\]
Under proper assumptions, $\sum_K\langle \widehat{u},v \rangle_{\partial K} = \langle \widehat{u}, \jump{v} \rangle_{\Gh}$, with energy setting
\[
u\in L^2\LRp{\Oh} \equiv L^2(\Omega), \quad v\in V=D(A^*_h), \quad
\widehat{u}\in \gamma(D(A)),
\]
where $D(A_h^*)$ is the broken graph space of the formal adjoint $A_h^*$, and $\gamma(D(A))$ the trace space of the graph space of operator $A$.\footnote{\tiny\bibentry{analysisDPG}}
}

\frame{
\frametitle{The canonical ``graph'' test norm}
Recall $\nor{u}_E \coloneqq \sup_{v\in V\setminus\{0\}}\frac{b(u,v)}{\nor{v}_V}$. Under the ultra-weak formulation, the trial norm 
\[
\nor{\LRp{u,\widehat{u}}}_U^2 \coloneqq \|u\|^2_{\L} + \|\widehat{u}\|_{\dag}^2
\] 
generates a test norm \emph{equivalent} to 
\[
\nor{v}_{V_{\rm opt}}^2 \coloneqq \|A_h^*v\|_{\L}^2 +\left(\sup_{\widehat{u}} \frac{\LRa{ \widehat{u},\jump{v} }_{\Gh}}{\|\widehat{u}\|}\right)^2.
\]
This norm is not localizable, so we instead substitute an $L^2$ term for the jump terms, giving us the (equivalent) \textcolor{red}{graph norm}\footnote{\tiny\bibentry{Bui-ThanhDemkowiczGhattas11b}}
\[
\nor{v}_V^2 \coloneqq \|A_h^*v\|_{\L}^2 + \nor{v}_{L^2}^2.
\]
\let\thefootnote\relax\footnotetext{\tiny{\dag: This norm has a marvelous technical definition for which this footnote is too small to contain.}}
}

\frame{
\frametitle{Global properties of test spaces: $L^2$-best approximations}

Under the ultra-weak formulation for $Au = f$.  
\[
b(\LRp{u,\uh},v) = \LRa{\widehat{u},\jump{v}}_{\Gh} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh},
\]
we recover the \textcolor{red}{$L^2$-projection} for \textcolor{red}{conforming} test functions that satisfy the adjoint equation:
\[
A_h^*v = \delta u, \quad \forall \delta u \in U_h. %v\in V^r = \LRc{v \in V(\Oh): \LRa{\uh,\jump{v}}_{\Gh} = 0, \quad \forall \LRp{0,\widehat{\delta u}} \in U}.
\]

For $\nor{v}_{\rm graph}^2 = \nor{A_h^*v}_{\L}^2 + \delta\nor{v}_{\L}^2$ as $\delta \rightarrow 0$; test functions solve
\[
\LRp{A_h^*v,A_h^*\delta v}_{L^2(K)} = (u,A_h^*\delta v)_{L^2(K)}.
\]
What can we say about \textcolor{red}{global properties} of local test functions?
}

\frame{
\frametitle{Global properties of DPG test spaces}
If we define $V^r_{{\rm opt}}$ as the optimal test space constructed using $V^r_p$, the \textcolor{red}{weakly conforming} space $\LRc{v \in V, \LRa{\widehat{\delta u},\jump{v}}_{\Gh} = 0, \quad \forall \LRp{0,\widehat{\delta u}} \in U_h}$,
\[
V^r_{{\rm opt}} = \LRc{ R_{V^r_p}^{-1}B \delta u, \quad \forall \delta u \in U_h},
\]
we have the following theorem\footnote{\tiny\bibentry{globalLocalDPG}}
\begin{theorem}
$V^r_{{\rm opt}} \subset V_{{\rm opt}}$.  Furthermore, the \textbf{field solutions} $u$ of $b(\LRp{u,\uh},v)=l(v)$ over $V_{{\rm opt}}$ and ${V}^r_{\rm opt}$ coincide.
\end{theorem}
\begin{itemize}
\item Locally determined test functions approximate global test functions.
\item Under a variant of the graph norm, \textcolor{red}{DPG approximates test functions which deliver $L^2$ optimality}.
\end{itemize}
}

%\frame{
%\frametitle{DPG as a non-conforming DG method on $V$}
%For DPG using the ultra-weak variational formulation
%\[
%b\left(\left(u, \widehat{u}\right),v\right) \coloneqq \LRa{\widehat{u},\jump{v}}_{\Gh} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh},
%\]
%under the graph test norm $\nor{v}_{\rm graph} = \nor{A_h^*v}_{\L} + \delta\nor{v}_{\L}$, as $\delta\rightarrow 0$, weakly conforming optimal test functions $\tilde{v}\in \tilde{V}$ satisfy 
%\[
%\LRp{A_h^*\tilde{v},A_h^*\tilde{\delta v}}_{\L} = (u,A_h^*\tilde{\delta v})_{\Oh}
%\]
%which yields a global \textcolor{red}{non-conforming least squares} method for $A_h^*v = u$.
%%\[
%%b\left(\left(u, \widehat{u}\right),\tilde{v}\right) = (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh}.
%%\]
%}

%\frame{%}

\subsection{A Robust DPG Method for Convection-Diffusion}
\begin{frame}{Outline}
  \tableofcontents[currentsection]
\end{frame}

\frame{
\frametitle{Ultra-weak formulation for convection-diffusion}
\vspace{-.25cm}
\begin{columns}
\begin{column}{.67\textwidth}
Convection-diffusion as a first order system is
\[
A \LRp{u,\sigma} \coloneqq 
\LRs {
\begin{array}{c}
\div (\beta u - \sigma) \\ \frac{1}{\epsilon}\sigma - \grad u
\end{array}} = \LRs{
\begin{array}{c}
f \\ 0
\end{array}
}.
\]
\end{column}
\begin{column}{.32\textwidth}
\begin{figure}[!h]
\centering
\includegraphics[width = 1.25in]{figs/hybridDGnodes.pdf}
\end{figure}
\end{column}
\end{columns}
\vspace{-.25cm}
The variational formulation is
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\grad_h\cdot \tau - \beta \cdot \grad_h
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad_h v\right)_{\Oh}&\\
  - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}&,
\end{align*}
where $\widehat{f}_n \coloneqq \beta_n u - \sigma_n$ and $\LRa{\widehat{f}_n,\jump{v} }_{\Gh}$ is defined 
\[
\LRa{\widehat{f}_n,\jump{v} }_{\Gh}\coloneqq \sum_K\int_{\partial K}{\rm sgn}(\vec{n}) \, \widehat{f}_n v.
\]
}

\frame{
\frametitle{Graph norm under convection-diffusion}
For convection-diffusion, the graph test norm is defined elementwise
\[
\|\left(v,\tau\right)\|_{V(K)}^2 = \| \div \tau - \beta \cdot \grad v
\|_{L^2(K)}^2 + \| \epsilon^{-1} \tau + \grad v \|_{L^2(K)}^2 +
\|v\|_{L^2(K)}^2.
\]
Problem with this test norm: approximability of test functions.\footnote{\tiny\bibentry{globalLocalDPG}}
\vspace{-.3cm}
\begin{figure}[!h]
\centering
\subfigure[Element $\tau$]{\includegraphics[width=1.1in]{figs/u_graph_tau_mag_K.pdf}}
\subfigure[Element $v$]{\includegraphics[width=1.1in]{figs/u_graph_v_K.pdf}}
\subfigure[Global $\tau$]{\includegraphics[width=1.1in]{figs/u_graph_tau_mag.pdf}}
\subfigure[Global $v$]{\includegraphics[width=1.1in]{figs/u_graph_v.pdf}}
\caption{$v$ and $\tau$ components of optimal test functions for $u = 1$ on $K$, for $\beta = (1,0)$ and $\epsilon = 0.01$. }
\label{fig:optTestBoundary}
\end{figure}
}
